What's hot and what's not:\\Windowing developer topic analysis

### 2009-04-10 lda-paper.tex
Added more content more edits, submitted to migod and holt

In our first exploratory pass we wanted to see if LDA could provide
interesting topics extracted from a real system. We took the
repository of MySQL 3.23, extracted the commits, grouped them by 30
day non-overlapping windows, and then applied LDA to each of these
windows. We asked LDA to make 20 topics per window and we then
carefully examined the top 10 most frequent words in that topic.  We
found there were common words across topic clusters, such as
\emph{diffs}, \emph{annotate} and \emph{history}, these words probably should
be viewed as stop words. There were notable transitional topics such
as in the first window the word \emph{bitkeeper} appears because MySQL
adopted Bitkeeper for their source control system yet in the following
windows there were no mentions of Bitkeeper. \emph{RENAME} also only
appeared once in the first window and never again. Per each topic we
tried to identify the purpose of the topic; to our surprise we found
that even with onlyy a little familiarity with the code base that
naming the topic was straightforward.

@@

The first visualization is a compressed view
(Figure\ref{fig:topicsmear} and Figure \ref{fig:zoomedsmear}) where we
try to sink the larger trends to the bottom.  Once the larger trends
are stacked we fill in the gaps with the smaller trends in the same
window, and then stack the smaller trends on top.  Although there is a
chance that gaps will be left due to non-optimal stacking in practice
there are lot of small trends (90\% of all trends contain 1 topic)
which fill in these gaps quickly.  Different instances of the same
trend share the same colour; apart from that, the color of trends is
randomly chosen and color similarity is not meaningful.
The compressed view is convienant because it shows all of the
information and the repeating continuous trends are easy to pick out,
although repeating discontinuous trends are harder to spot.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{histogram-cropped-scaled}
  \caption{Top part of trend histogram, ordered by topic occurance for MaxDB 7.500. X axis determines the number of continuous months of a trend. Trends are ranked  by the number of topics that a trend contains in descending order.}       
  \label{fig:histogram}
\end{figure}

@@

We extracted the repositories of: PostGreSQL, MaxDB and
Firebird.

@@

\subsection{Postgresql}
%   3.   mysql

@@

The plots we produced of MaxDB 7.500 were unlike those of the other
systems, as there was a period where no real development occurred and
thus there were no topics or trends whatsoever. We have two periods to
evaluate for MaxDB 7.500, the first period from June 2004 to jan 2005,
and the second period from June 2005 to June 2006.

@@

The query optimizer was a topic which recurred during
MaxDB's development. In our plots, topics which mention optimizer
occur four times yet in the large static plot (Figure
\ref{fig:statictopics}) it is not in any of the topics. A query
optimizer is rather important to an DBMS, but as we've shown it just
doesn't pop up as a topic on its own. We tried to remove words to see
if we could get optimizer topic out, after removing stop words and
then two of the most common words the global analysis finally had a
topic with optimizer in its topic ten words. This shows that optimizer
was important but it had been obscured, while it would have been
noticed using a more locally relevant approach that we propose.

@@

UNICODE

DEC/OSF1

ODBC

crash

@@

SQLCLI
%V75

SYSDD.cpnix
backup

select
make
memory
view

debug

@@

If we approach the difference via common tokens we can see that common
token tend to dominate both results, at least in the case of MaxDB
7.500 our local topic approach netted us topics which contained these
relatively important and common words: \emph{UNICODE},
\emph{DEC/OSF1}, \emph{ODBC}, \emph{crash}, \emph{SQLCLI},
\emph{SYSDD.cpnix}, \emph{backup}, \emph{select}, \emph{make},
\emph{memory}, \emph{view}, and finally \emph{debug}. None of these
tokens were part of the static topic analysis topics but they were
part of 494 commits (5\% of the entire system) to MaxDB 7.500.

@@

XXX Numbers and more argument

@@

XXX summarize our position

XXX Add more detail here

We presented multiple visualizations that focused on different
aspects of the data: temporality of trends, trend size, and a
condensed overall trend view. The condensed trend view shows much more
information than the views that global analysis could show.

>>>>

In our first exploratory pass we wanted to see if LDA could provide
interesting topics extracted from a real system. We took the
repository of MySQL 3.23, extracted the commits, grouped them by 30
day non-overlapping windows, and then applied LDA to each of these
windows. We asked LDA to make 20 topics per window and we then
carefully examined the top 10 most frequent words in that topic.  We
found there were common words across topic clusters, such as
\emph{diffs}, \emph{annotate} and \emph{history}, these words probably should
be viewed as stop words. There were notable transitional topics such
as in the first window the word \emph{bitkeeper} appears because MySQL
adopted Bitkeeper for their source control system yet in the following
windows there were no mentions of Bitkeeper. \emph{RENAME} also only
appeared once in the first window and never again. Per each topic we
tried to identify the purpose of the topic; to our surprise we found
that even with only a little familiarity with the code base that
naming the topic was straightforward.

@@

The first visualization is a compressed view
(Figure\ref{fig:topicsmear} and Figure \ref{fig:zoomedsmear}) where we
try to sink the larger trends to the bottom.  Once the larger trends
are stacked we fill in the gaps with the smaller trends in the same
window, and then stack the smaller trends on top.  Although there is a
chance that gaps will be left due to non-optimal stacking in practice
there are lot of small trends (90\% of all trends contain 1 topic)
which fill in these gaps quickly.  Different instances of the same
trend share the same color; apart from that, the color of trends is
randomly chosen and color similarity is not meaningful.
The compressed view is convenient because it shows all of the
information and the repeating continuous trends are easy to pick out,
although repeating discontinuous trends are harder to spot.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{histogram-cropped-scaled}
  \caption{Top part of trend histogram, ordered by topic occurrence for MaxDB 7.500. X axis determines the number of continuous months of a trend. Trends are ranked  by the number of topics that a trend contains in descending order.}       
  \label{fig:histogram}
\end{figure}

@@

We extracted the repositories of: PosgreSQL, MaxDB and
Firebird.

@@

\subsection{PostgreSQL}
%   3.   mysql

@@

The plots we produced of MaxDB 7.500 were unlike those of the other
systems, as there was a period where no real development occurred and
thus there were no topics or trends whatsoever. We have two periods to
evaluate for MaxDB 7.500, the first period from June 2004 to January 2005,
and the second period from June 2005 to June 2006.

@@

The query optimizer was a topic which recurred during
MaxDB's development. In our plots, topics which mention optimizer
occur four times yet in the large static plot (Figure
\ref{fig:statictopics}) it is not in any of the topics. A query
optimizer is rather important to an DBMS, but as we have shown it just
does not pop up as a topic on its own. We tried to remove words to see
if we could get optimizer topic out, after removing stop words and
then two of the most common words the global analysis finally had a
topic with optimizer in its topic ten words. This shows that optimizer
was important but it had been obscured, while it would have been
noticed using a more locally relevant approach that we propose.

We noticed that PHP was in a third of the commits that Perl was in
(not the same commits) in as many commits as Perl yet PHP did not
appear in the global static topic analysis topics while Perl did. Our
local topic analysis mentioned PHP in 5 different topics, yet
mentioned Perl in 4 different topics. Perhaps the Perl mentions were
not as independent topics as the PHP topics were.

@@

% UNICODE

% DEC/OSF1

% ODBC

% crash

@@

% SQLCLI
% %V75

% SYSDD.cpnix
% backup

% select
% make
% memory
% view

% debug

@@

@@

If we approach the difference via common tokens we can see that common
token tend to dominate both results, at least in the case of MaxDB
7.500 our local topic approach netted us topics which contained these
relatively important and common words: \emph{UNICODE},
\emph{DEC/OSF1}, \emph{ODBC}, \emph{crash}, \emph{SQLCLI},
\emph{SYSDD.cpnix}, \emph{backup}, \emph{select}, \emph{make},
\emph{memory}, \emph{view}, and finally \emph{debug}. None of these
tokens were part of the static topic analysis topics, but there were
part of 566 commits (6\% of the entire system) to MaxDB 7.500 which
contained these. These tokens were part of 87 out of 500 of our
locally generated topics (ignoring similar topics).

%XXX Numbers and more argument

@@

We demonstrated the application of windowed topic analysis, topic
analysis applied to commit messages during periods of
development. This approach differed from previous work, as previous
work would apply topic analysis globally and then analyze related
documents. Instead we show that many topics that exist locally are
relevant and interesting yet would often not be included in global
topic analysis. We handle the case of recurring topics by using a
topic similarity measure which looks for topics which recur and mutate
throughout the development of the software project.

The value of windowed topic analysis was demonstrated on a case study
of three database systems with a particular focus on MaxDB 7.500. We
showed that static topic analysis missed important database concepts
such as ODBC.

%XXX summarize our position

%XXX Add more detail here

We presented multiple visualizations that focused on different aspects
of the data: temporality of trends, trend size, and a condensed
overall trend view. The condensed trend view shows much more
information than the views that global analysis could show, it
indicates how focused a period is by the total number of topics as
well shows topics by similarity so one can track trends across time.
Our histogram trend-view immediately hi-lights and measures the size of
trends while our time-line trend view shows how a topic reoccurs over
time. These visualizations help us get an over feeling about the
common and unique topics that developers focus on. If implemented
properly a user can easily zoom and query a topic.
